---
title: "Optimal Control Results"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# markdown settings
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# required packages
library(ggplot2)
library(deSolve)
library(reshape2)
library(tidyverse)
library(cowplot)
library(pracma)
library(RColorBrewer)

# to paralellize
library(doParallel)
library(foreach)
registerDoParallel(detectCores() - 2) # update this if you want to use more cores
```

```{r oc_setup}
# load optimal control files
source("implementation/optimal_control_functions.R")
# source("run_optimal_control.R")

# initial guesses - controls
guess_v1 <- rep(0, length(times))
guess_v2 <- rep(0, length(times))
guess_u1 <- rep(0, length(times))
guess_u2 <- rep(0, length(times))

# setup baseline optimal control parameters
tol <- .01 # tolerance parameter for optimization
oc_params <- c(
  b1 = 1, b2 = 1, # cost of cases
  C1 = 0.125, C2 = 0.125, # cost of vaccinations
  epsilon1 = 10000, epsilon2 = 10000, # non-linearity for vacc
  D1 = 0.0125, D2 = 0.0125, # cost of sanitation
  eta1 = 1000, eta2 = 1000
) # non-linearity for sanitation
```

```{r model_setup}
# Load setup
cholera_setup <- setup_model("cholera")
ebola_setup <- setup_model("ebola")
# Initial conditions
cholera_IC <- cholera_setup$init_x
ebola_IC <- ebola_setup$init_x
# Parameters
cholera_params <- cholera_setup$params
ebola_params <- ebola_setup$params
```

```{r plot_setup}
# define color palettes used in plots
patch_colors2 <- c("#DE2D26", "#3182BD")
cost_colors6 <- c(rev(brewer.pal(3, "Reds")), rev(brewer.pal(3, "Blues")))

# Define functions for plotting

# Plot trajectories with NO CONTROLS
no_control_plot_function <- function(df, model) {
  if (model == "ebola") {
    level_codes <- c("S", "E", "I", "H", "D", "R")
    compartment_names <- c("Susceptible", "Exposed", "Infected", "Hospitalized",
                           "Dead", "Recovered")
  } else {
    level_codes <- c("S", "I", "R", "W")
    compartment_names <- c("Susceptible", "Infected", "Recovered", "Water")
  }
  names(compartment_names) <- level_codes
  
  df %>%
    select(-v1, -v2, -u1, -u2) %>%
    # melt from wide to long format
    melt(c("time")) %>%
    # use variable names to distinguish compartment and patch
    mutate(
      compartment = substr(variable, 1, 1),
      compartment = factor(compartment, levels = level_codes),
      patch = substr(variable, 2, 2)
    ) %>%
    # plot time series
    ggplot(aes(x = time, y = value, color = patch)) +
    geom_line(size = 1) +
    facet_wrap(vars(compartment), scales = "free", 
               labeller = labeller(variable = compartment_names)) +
    scale_color_manual(values = patch_colors2) +
    theme_bw() +
    theme(
      axis.title.y = element_blank(),
      legend.position = "bottom",
      panel.grid = element_blank()
    )
  
}

# Plot trajectories of each compartment including controls
trajectory_plot_function <- function(df, model) {
  if (model == "ebola") {
    level_codes <- c("S", "E", "I", "H", "D", "R", "v", "u")
    compartment_names <- c("Susceptible", "Exposed", "Infected", "Hospitalized", "Dead", "Recovered", "Vaccinated", "TBD")
  } else {
    level_codes <- c("S", "I", "R", "W", "v", "u")
    compartment_names <- c("Susceptible", "Infected", "Recovered", "Water", "Vaccinated", "Sanitation")
  }
  names(compartment_names) <- level_codes
  
  plot <- df %>%
    select(-c(m1, m2, u1_max, u2_max, v1_max, v2_max, test_case,
              j_case1, j_case2)) %>% 
    melt(c("control_type", "time")) %>%
    mutate(
      patch = substr(variable, 2, 2),
      variable = substr(variable, 1, 1)
    ) %>%
    filter(control_type %in% c("uniform", "unique")) %>%
    mutate(
      plot_var = ifelse(control_type == "unique", paste("patch", patch), "uniform"),
      variable = factor(variable, levels = level_codes)
    ) %>%
    filter(!is.na(variable)) %>%
    ggplot(aes(x = time, y = value, color = patch, linetype = control_type)) +
    geom_line(size = 1) +
    facet_wrap(vars(variable), scales = "free", labeller = labeller(variable = compartment_names)) +
    scale_color_manual(values = patch_colors2) +
    scale_linetype_manual(values = c("dotted", "solid")) +
    theme_bw() +
    ggtitle("Trajectory comparisons") +
    theme(
      legend.position = "bottom",
      panel.grid = element_blank()
    )
  return(plot)
}

# Plot bar graph of costs
cost_plot_function <- function(df, model) {
  if (model == "ebola") {
    cost_levels <- c(
      "j_case1", "j_vacc1", "j_sani1", "j_case2", "j_vacc2",
      "j_sani2"
    )
    cost_names <- c(
      "P1: cases", "P1: vaccination", "P1: ?",
      "P2: cases", "P2: vaccination", "P2: ?"
    )
    clrs <- cost_colors6[c(1:2,4:5)]
  } else {
    cost_levels <- c(
      "j_case1", "j_vacc1", "j_sani1", "j_case2", "j_vacc2",
      "j_sani2"
    )
    cost_names <- c(
      "P1: cases", "P1: vaccination", "P1: sanitation",
      "P2: cases", "P2: vaccination", "P2: sanitation"
    )
    clrs <- cost_colors6
  }
  names(cost_names) <- cost_levels
  
  plt <- df %>%
    select(test_case, control_type, contains("j")) %>% 
    melt(c("control_type", "test_case")) %>%
    filter(variable != "j") %>%
    mutate(variable = factor(variable, levels = cost_levels)) %>%
    unique()
  ggplot(data = plt, aes(x = control_type)) +
    geom_bar(aes(y = value, fill = variable),
             stat = "identity", position = "stack"
    ) +
    geom_text(aes(y = value, label = round(value), group = variable),
              position = position_stack(vjust = 0.5)
    ) +
    geom_text(
      data = plt %>% group_by(control_type) %>% summarise(value = sum(value)),
      aes(y = value, label = paste0("Total cost: ", round(value))),
      vjust = 0
    ) +
    guides(fill = guide_legend(nrow = length(clrs)/2)) +
    scale_fill_manual(values = clrs, labels = cost_names) +
    labs(y = "cost") +
    scale_x_discrete(expand = c(0, 0)) +
    theme_bw() +
    ggtitle("Cost breakdown") +
    theme(
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.grid = element_blank()
    )
}


# Plot costs of uniform case relative to unique
relcost_plot_function <- function(df, model) { 
  if (model == "cholera") {
    change_vars <- c("j", "j_case1", "j_case2", "j_vacc1", "j_vacc2", "j_sani1", "j_sani2")
    names(change_vars) <- c("Total cost", "Cases in patch 1", "Cases in patch 2", 
                            "Vaccinations in patch 1", "Vaccinations in patch 2",
                            "Sanitation in patch 1", "Sanitation in patch 2")
  } else {
    change_vars <- c("j", "j_case1", "j_case2", "j_vacc1", "j_vacc2")
    names(change_vars) <- c("Total cost", "Cases in patch 1", "Cases in patch 2", 
                            "Vaccinations in patch 1", "Vaccinations in patch 2")
  }
  
  rel_costs <- df %>%
    select(control_type, change_vars) %>%
    unique() %>%
    melt(id.var = "control_type") %>% 
    dcast(variable ~ control_type) %>%
    mutate(rel_change = (unique - uniform)/uniform,
           variable = factor(variable, levels = names(change_vars)))
  
  # rel_costs <- df %>%
  #   select(test_case, control_type, contains("j")) %>% 
  #   mutate(total_cases1 = j_case1 / params$b1) %>%
  #   mutate(total_cases2 = j_case2 / params$b2) %>%
  #   mutate(total_vacc1 = j_vacc1 / params$C1) %>%
  #   mutate(total_vacc2 = j_vacc2 / params$C2) %>%
  #   # mutate(total_sani1 = j_sani1 / params$D1) %>%
  #   # mutate(total_sani2 = j_sani2 / params$D2) %>%
  #   mutate(case_diff = total_cases1 - total_cases2) %>%
  #   mutate(vacc_diff = total_vacc1 - total_vacc2) %>%
  #   # mutate(sani_diff = total_sani1 - total_sani2) %>%
  #   mutate(`Total cost` = 100 * (j - j[control_type == "uniform"]) / j[control_type == "uniform"]) %>%
  #   mutate(`Cases in patch 1` = 100 * (total_cases1 - total_cases1[control_type == "uniform"]) /
  #            total_cases1[control_type == "uniform"]) %>%
  #   mutate(`Cases in patch 2` = 100 * (total_cases2 - total_cases2[control_type == "uniform"]) /
  #            total_cases2[control_type == "uniform"]) %>%
  #   mutate(`Vaccinations in patch 1` = 100 * (total_vacc1 - total_vacc1[control_type == "uniform"]) /
  #            total_vacc1[control_type == "uniform"]) %>%
  #   mutate(`Vaccinations in patch 2` = 100 * (total_vacc2 - total_vacc2[control_type == "uniform"]) /
  #            total_vacc2[control_type == "uniform"]) %>%
  #   filter(control_type == "unique")
  
  rel_costs %>%
    ggplot(aes(x = variable, y = rel_change)) +
    geom_col() +
    ggtitle("Relative changes due to switching from Uniform to Unique") +
    labs(y = "% change", x = "") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

# Cholera {.tabset}
```{r cholera-plot_setup}
# define labels to make plots more interpretable
# names of each compartment/control
compartment_names <- c(
  "Susceptible", "Infected", "Recovered", "Water",
  "vaccination", "sanitation"
)
names(compartment_names) <- c("S", "I", "R", "W", "v", "u")
# names of each aspect of cost function
cost_names <- c(
  "P1: cases", "P1: vaccination", "P1: sanitation",
  "P2: cases", "P2: vaccination", "P2: sanitation"
)
names(cost_names) <- c(
  "j_case1", "j_vacc1", "j_sani1",
  "j_case2", "j_vacc2", "j_sani2"
)


# make parameter tables look nice
# KD: this is a temporary fix to make things look nice.
#     I think the way I reconfigured on my branch will allow us to avoid doing this
params_table <- matrix(as.list(cholera_params), 1, length(as.list(cholera_params)))
colnames(params_table) <- names(as.list(cholera_params))
params_table <- as.data.frame(params_table)

IC_table <- as.data.frame(t(cholera_IC))

# Calculate R0
choleraR0 <- R0_cholera(cholera_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```

```{r cholera-calculate}
# setup all parameter sets to test
test_params <- expand.grid(
  control_type = c("unique", "uniform"),
  m1 = c(0, .05, .1),
  m2 = c(0, .05, .1),
  u1_max = c(0, 0.4),
  u2_max = c(0, 0.4),
  v1_max = c(0, 0.015),
  v2_max = c(0, 0.015)
) %>% 
  as.data.frame() %>% 
  filter_(~u1_max == u2_max) %>% 
  filter_(~v1_max == v2_max)

# calculate all trajectories and costs
cholera_optim_full <- mult_oc_optim("cholera", test_params)
```

Here we keep the results of various numerical experiments of optimal control outcomes in the Cholera model. Specifically, we focus on the difference between the optimal strategy, when control is customized to each patch versus the "one size fits all strategy" where both patches are forced to be given the same control.

## Experiment 0 (no movement between patches) {.tabset}

```{r cholera-exp0}
cholera_exp0_df <- filter(cholera_optim_full,
                       m1 == 0,
                       m2 == 0)
# Vaccination only
exper_v_only <- filter(cholera_exp0_df,
                       u1_max == 0,
                       v1_max == params_cholera$v1_max
)
# Sanitation only
exper_u_only <- filter(cholera_exp0_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == 0
)
# Both controls
exper_both <- filter(cholera_exp0_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == params_cholera$v1_max
)
```

### Parameters 
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : gamma2), format="markdown")`
`r knitr::kable(params_table %>% select(delta1 : rho2), format="markdown")`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : control_type), format="markdown")`
`r knitr::kable(params_table %>% select(v1_min : u2_max), format="markdown")`

With these parameters, the system has an R0 of about `r choleraR0`.

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
```{r cholera-exp0-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("cholera", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "cholera")
```


### Vaccination Only
We consider the case with only vaccination (no sanitation).

Controls start at day `r response_time`.

```{r cholera-exp0-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "cholera")
```

```{r cholera-exp0-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "cholera")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "cholera")
```

### Sanitation Only
We also consider the case with only sanitation (no vaccination).

Controls start at day `r response_time``.

```{r cholera-exp0-sani_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "cholera")
```

```{r cholera-exp0-sani_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")

# plot costs
cost_plot_function(exper_u_only, "cholera")
```

### Both Vaccination and Sanitation

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r cholera-exp0-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "cholera")
```

```{r cholera-exp0-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")
# plot costs
cost_plot_function(exper_both, "cholera")
```

## Experiment 1 (symmetric, m1=m2=0.05) {.tabset}
```{r cholera-exp1}
cholera_exp1_df <- filter(cholera_optim_full,
                       m1 == 0.05,
                       m2 == 0.05)

# Vaccination only
exper_v_only <- filter(cholera_exp1_df,
                       u1_max == 0,
                       v1_max == params_cholera$v1_max
)
# Sanitation only
exper_u_only <- filter(cholera_exp1_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == 0
)
# Both controls
exper_both <- filter(cholera_exp1_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == params_cholera$v1_max
)
```
### Parameters 
```{r cholera-exp1-parameters}
exp1_params <- cholera_params
exp1_params$m1 <- 0.05
exp1_params$m2 <- 0.05

choleraR0 <- R0_cholera(exp1_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(exp1_params %>% select(mu1 : gamma2), format="markdown")`
`r knitr::kable(exp1_params %>% select(delta1 : rho2), format="markdown")`

Initial conditions: **need to re-calculate initial conditions with movement** `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(exp1_params %>% select(b1 : control_type), format="markdown")`
`r knitr::kable(exp1_params %>% select(v1_min : u2_max), format="markdown")`

With these parameters, the system has an R0 of about `r choleraR0`.

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r cholera-exp1-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("cholera", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "cholera")
```

### Vaccination Only
We consider the case with only vaccination (no sanitation).

Controls start at day `r response_time`.

```{r cholera-exp1-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "cholera")
```

```{r cholera-exp1-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "cholera")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "cholera")
```

### Sanitation Only
We also consider the case with only sanitation (no vaccination).

Controls start at day `r response_time``.

```{r cholera-exp1-sani_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "cholera")
```

```{r cholera-exp1-sani_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")

# plot costs
cost_plot_function(exper_u_only, "cholera")
```

### Both Vaccination and Sanitation

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r cholera-exp1-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "cholera")
```

```{r cholera-exp1-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")
# plot costs
cost_plot_function(exper_both, "cholera")
```

## Experiment 2 (symmetric movement, m1=m2=0.1) {.tabset}
```{r cholera-exp2}
cholera_exp2_df <- filter(cholera_optim_full,
                       m1 == 0.1,
                       m2 == 0.1)

# Vaccination only
exper_v_only <- filter(cholera_exp2_df,
                       u1_max == 0,
                       v1_max == params_cholera$v1_max
)
# Sanitation only
exper_u_only <- filter(cholera_exp2_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == 0
)
# Both controls
exper_both <- filter(cholera_exp2_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == params_cholera$v1_max
)
```
### Parameters 
```{r cholera-exp2-parameters}
exp2_params <- cholera_params
exp2_params$m1 <- 0.1
exp2_params$m2 <- 0.1

choleraR0 <- R0_cholera(exp2_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(exp2_params %>% select(mu1 : gamma2), format="markdown")`
`r knitr::kable(exp2_params %>% select(delta1 : rho2), format="markdown")`

Initial conditions: **need to re-calculate initial conditions with movement ** `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(exp2_params %>% select(b1 : control_type), format="markdown")`
`r knitr::kable(exp2_params %>% select(v1_min : u2_max), format="markdown")`

With these parameters, the system has an R0 of about `r choleraR0`.

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r cholera-exp2-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("cholera", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "cholera")
```

### Vaccination Only
We consider the case with only vaccination (no sanitation).

Controls start at day `r response_time`.

```{r cholera-exp2-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "cholera")
```

```{r cholera-exp2-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "cholera")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "cholera")
```

### Sanitation Only
We also consider the case with only sanitation (no vaccination).

Controls start at day `r response_time``.

```{r cholera-exp2-sani_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "cholera")
```

```{r cholera-exp2-sani_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")

# plot costs
cost_plot_function(exper_u_only, "cholera")
```

### Both Vaccination and Sanitation

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r cholera-exp2-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "cholera")
```

```{r cholera-exp2-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")
# plot costs
cost_plot_function(exper_both, "cholera")
```

## Experiment 3 (asymmetric movement, m1=0.05, m2=0.1) {.tabset}
```{r cholera-exp3}
cholera_exp3_df <- filter(cholera_optim_full,
                       m1 == 0.05,
                       m2 == 0.1)

# Vaccination only
exper_v_only <- filter(cholera_exp3_df,
                       u1_max == 0,
                       v1_max == params_cholera$v1_max
)
# Sanitation only
exper_u_only <- filter(cholera_exp3_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == 0
)
# Both controls
exper_both <- filter(cholera_exp3_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == params_cholera$v1_max
)
```
### Parameters 
```{r cholera-exp3-parameters}
exp3_params <- cholera_params
exp3_params$m1 <- 0.05
exp3_params$m2 <- 0.1

choleraR0 <- R0_cholera(exp3_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(exp3_params %>% select(mu1 : gamma2), format="markdown")`
`r knitr::kable(exp3_params %>% select(delta1 : rho2), format="markdown")`

Initial conditions: **need to re-calculate initial conditions with movement ** `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(exp3_params %>% select(b1 : control_type), format="markdown")`
`r knitr::kable(exp3_params %>% select(v1_min : u2_max), format="markdown")`

With these parameters, the system has an R0 of about `r choleraR0`.

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r cholera-exp3-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("cholera", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "cholera")
```

### Vaccination Only
We consider the case with only vaccination (no sanitation).

Controls start at day `r response_time`.

```{r cholera-exp3-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "cholera")
```

```{r cholera-exp3-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "cholera")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "cholera")
```

### Sanitation Only
We also consider the case with only sanitation (no vaccination).

Controls start at day `r response_time``.

```{r cholera-exp3-sani_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "cholera")
```

```{r cholera-exp3-sani_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")

# plot costs
cost_plot_function(exper_u_only, "cholera")
```

### Both Vaccination and Sanitation

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r cholera-exp3-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "cholera")
```

```{r cholera-exp3-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")
# plot costs
cost_plot_function(exper_both, "cholera")
```

## Experiment 4 (asymmetric movement, m1=0.1, m2 =0.05) {.tabset}
```{r cholera-exp4}
cholera_exp4_df <- filter(cholera_optim_full,
                       m1 == 0.1,
                       m2 == 0.05)

# Vaccination only
exper_v_only <- filter(cholera_exp4_df,
                       u1_max == 0,
                       v1_max == params_cholera$v1_max
)
# Sanitation only
exper_u_only <- filter(cholera_exp4_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == 0
)
# Both controls
exper_both <- filter(cholera_exp4_df,
                       u1_max == params_cholera$u1_max,
                       v1_max == params_cholera$v1_max
)
```
### Parameters 
```{r cholera-exp4-parameters}
exp4_params <- cholera_params
exp4_params$m1 <- 0.05
exp4_params$m2 <- 0.1

choleraR0 <- R0_cholera(exp4_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(exp4_params %>% select(mu1 : gamma2), format="markdown")`
`r knitr::kable(exp4_params %>% select(delta1 : rho2), format="markdown")`

Initial conditions: **need to re-calculate initial conditions with movement ** `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(exp4_params %>% select(b1 : control_type), format="markdown")`
`r knitr::kable(exp4_params %>% select(v1_min : u2_max), format="markdown")`

With these parameters, the system has an R0 of about `r choleraR0`.

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r cholera-exp4-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("cholera", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "cholera")
```

### Vaccination Only
We consider the case with only vaccination (no sanitation).

Controls start at day `r response_time`.

```{r cholera-exp4-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "cholera")
```

```{r cholera-exp4-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "cholera")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "cholera")
```

### Sanitation Only
We also consider the case with only sanitation (no vaccination).

Controls start at day `r response_time``.

```{r cholera-exp4-sani_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "cholera")
```

```{r cholera-exp4-sani_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")

# plot costs
cost_plot_function(exper_u_only, "cholera")
```

### Both Vaccination and Sanitation

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r cholera-exp4-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "cholera")
```

```{r cholera-exp4-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "cholera")
# plot costs
cost_plot_function(exper_both, "cholera")
```

# Ebola {.tabset}

```{r ebola-plot_setup}
# define labels to make plots more interpretable
# names of each compartment/control
compartment_names <- c(
  "Susceptible", "Exposed", "Infected", "Hospitalized", "Dead", "Recovered",
  "Vaccination", "TBD"
)
names(compartment_names) <- c("S", "E", "I", "H", "D", "R", "v", "u")
# names of each aspect of cost function
cost_names <- c(
  "P1: cases", "P1: vaccination", "P1: TBD",
  "P2: cases", "P2: vaccination", "P2: TBD"
)
names(cost_names) <- c(
  "j_case1", "j_vacc1", "j_sani1",
  "j_case2", "j_vacc2", "j_sani2"
)

# define color palettes for plots
patch_colors2 <- c("#DE2D26", "#3182BD")
cost_colors6 <- c(rev(brewer.pal(3, "Reds")), rev(brewer.pal(3, "Blues")))


# make parameter tables look nice
# KD: this is a temporary fix to make things look nice.
#     I think the way I reconfigured on my branch will allow us to avoid doing this
params_table <- matrix(as.list(ebola_params), 1, length(as.list(ebola_params)))
colnames(params_table) <- names(as.list(ebola_params))
params_table <- as.data.frame(ebola_params)

IC_table <- as.data.frame(t(ebola_IC))
# Calculate R0
#ebolaR0 <- R0_ebola(ebola_params, sum(IC_table[c(1, 3, 5)]), sum(IC_table[c(2, 4, 6)]))
```

```{r ebola-calculate}
# setup all parameter sets to test
test_params <- expand.grid(
  control_type = c("unique", "uniform"),
  m1 = c(0, .05, .1),
  m2 = c(0, .05, .1),
  u1_max = c(0, 0.4),
  u2_max = c(0, 0.4),
  v1_max = c(0, 0.015),
  v2_max = c(0, 0.015)
) %>% 
  as.data.frame() %>% 
  filter_(~u1_max == u2_max) %>% 
  filter_(~v1_max == v2_max)

# calculate all trajectories and costs
ebola_optim_full <- mult_oc_optim("ebola", test_params)
```
We show similar results instead employing  the Ebola model.

## Experiment 0 (no movement between patches) {.tabset}
```{r ebola-exp0}
ebola_exp0_df <- filter(ebola_optim_full,
                       m1 == 0,
                       m2 == 0)
# Vaccination only
exper_v_only <- filter(ebola_exp0_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
# TBD only
exper_u_only <- filter(ebola_exp0_df,
                       u1_max == 0,
                       v1_max == 0
)
# Both controls
exper_both <- filter(ebola_exp0_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
```
### Parameters
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : n1), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(mu2 : n2), format="markdown", align = 'l')`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : v2), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(v1_min : N2), format="markdown", align = 'l')`

*Need to add Ebola R0 calculation*

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r ebola-exp0-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("ebola", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "ebola")
```

### Vaccination Only
We consider the case with only vaccination (no TBD).

Controls start at day `r response_time`.

```{r ebola-exp0-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "ebola")
```

```{r ebola-exp0-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "ebola")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "ebola")
```

### TBD Only
We also consider the case with only TBD (no vaccination).

Controls start at day `r response_time``.

```{r ebola-exp0-TBD_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "ebola")
```

```{r ebola-exp0-TBD_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")

# plot costs
cost_plot_function(exper_u_only, "ebola")
```

### Both Vaccination and TBD

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r ebola-exp0-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "ebola")
```

```{r ebola-exp0-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")
# plot costs
cost_plot_function(exper_both, "ebola")
```

## Experiment 1 (symmetric, m1=m2=0.05) {.tabset}
```{r ebola-exp1}
ebola_exp1_df <- filter(ebola_optim_full,
                       m1 == 0.05,
                       m2 == 0.05)
# Vaccination only
exper_v_only <- filter(ebola_exp1_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
# TBD only
exper_u_only <- filter(ebola_exp1_df,
                       u1_max == 0,
                       v1_max == 0
)
# Both controls
exper_both <- filter(ebola_exp1_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
```
### Parameters
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : n1), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(mu2 : n2), format="markdown", align = 'l')`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : v2), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(v1_min : N2), format="markdown", align = 'l')`

*Need to add Ebola R0 calculation*

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r ebola-exp1-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("ebola", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "ebola")
```

### Vaccination Only
We consider the case with only vaccination (no TBD).

Controls start at day `r response_time`.

```{r ebola-exp1-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "ebola")
```

```{r ebola-exp1-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "ebola")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "ebola")
```

### TBD Only
We also consider the case with only TBD (no vaccination).

Controls start at day `r response_time``.

```{r ebola-exp1-TBD_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "ebola")
```

```{r ebola-exp1-TBD_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")

# plot costs
cost_plot_function(exper_u_only, "ebola")
```

### Both Vaccination and TBD

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r ebola-exp1-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "ebola")
```

```{r ebola-exp1-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")
# plot costs
cost_plot_function(exper_both, "ebola")
```

## Experiment 2 (symmetric, m1=m2=0.1) {.tabset}
```{r ebola-exp2}
ebola_exp2_df <- filter(ebola_optim_full,
                       m1 == 0.1,
                       m2 == 0.1)
# Vaccination only
exper_v_only <- filter(ebola_exp2_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
# TBD only
exper_u_only <- filter(ebola_exp2_df,
                       u1_max == 0,
                       v1_max == 0
)
# Both controls
exper_both <- filter(ebola_exp2_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
```
### Parameters
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : n1), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(mu2 : n2), format="markdown", align = 'l')`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : v2), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(v1_min : N2), format="markdown", align = 'l')`

*Need to add Ebola R0 calculation*

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r ebola-exp2-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("ebola", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "ebola")
```

### Vaccination Only
We consider the case with only vaccination (no TBD).

Controls start at day `r response_time`.

```{r ebola-exp2-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "ebola")
```

```{r ebola-exp2-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "ebola")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "ebola")
```

### TBD Only
We also consider the case with only TBD (no vaccination).

Controls start at day `r response_time``.

```{r ebola-exp2-TBD_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "ebola")
```

```{r ebola-exp2-TBD_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")

# plot costs
cost_plot_function(exper_u_only, "ebola")
```

### Both Vaccination and TBD

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r ebola-exp2-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "ebola")
```

```{r ebola-exp2-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")
# plot costs
cost_plot_function(exper_both, "ebola")
```


## Experiment 3 (asymmetric, m1=.05, m2=0.1) {.tabset}
```{r ebola-exp3}
ebola_exp3_df <- filter(ebola_optim_full,
                       m1 == 0.05,
                       m2 == 0.1)
# Vaccination only
exper_v_only <- filter(ebola_exp3_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
# TBD only
exper_u_only <- filter(ebola_exp3_df,
                       u1_max == 0,
                       v1_max == 0
)
# Both controls
exper_both <- filter(ebola_exp3_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
```
### Parameters
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : n1), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(mu2 : n2), format="markdown", align = 'l')`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : v2), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(v1_min : N2), format="markdown", align = 'l')`

*Need to add Ebola R0 calculation*

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r ebola-exp3-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("ebola", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "ebola")
```

### Vaccination Only
We consider the case with only vaccination (no TBD).

Controls start at day `r response_time`.

```{r ebola-exp3-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "ebola")
```

```{r ebola-exp3-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "ebola")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "ebola")
```

### TBD Only
We also consider the case with only TBD (no vaccination).

Controls start at day `r response_time``.

```{r ebola-exp3-TBD_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "ebola")
```

```{r ebola-exp3-TBD_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")

# plot costs
cost_plot_function(exper_u_only, "ebola")
```

### Both Vaccination and TBD

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r ebola-exp3-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "ebola")
```

```{r ebola-exp3-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")
# plot costs
cost_plot_function(exper_both, "ebola")
```



## Experiment 4 (asymmetric, m1=.1, m2=0.05) {.tabset}
```{r ebola-exp4}
ebola_exp4_df <- filter(ebola_optim_full,
                       m1 == 0.1,
                       m2 == 0.05)
# Vaccination only
exper_v_only <- filter(ebola_exp4_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
# TBD only
exper_u_only <- filter(ebola_exp4_df,
                       u1_max == 0,
                       v1_max == 0
)
# Both controls
exper_both <- filter(ebola_exp4_df,
                       u1_max == 0,
                       v1_max == params_ebola$v1_max
)
```
### Parameters
The following experiments use these  parameters and initial conditions.

Parameters: `r knitr::kable(params_table %>% select(mu1 : n1), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(mu2 : n2), format="markdown", align = 'l')`

Initial conditions: `r knitr::kable(IC_table, format =  "markdown", digits = 0, align = 'l')`

Optimal control parameters: `r knitr::kable(params_table %>% select(b1 : v2), format="markdown", align = 'l')`
`r knitr::kable(params_table %>% select(v1_min : N2), format="markdown", align = 'l')`

*Need to add Ebola R0 calculation*

### No control
Time 0 corresponds to `r response_time` days since the start of the outbreak.
**This is currently incorrect. It’s not taking into account movement**
```{r ebola-exp4-no-control, fig.show="hold", out.width = "100%"}
# read in csv with solution to ode without control
uncontrolled_epi <- run_no_optim("ebola", "none")$trajectories # KD: needs to be changed to allow varying parameters like movement

# plot the time series in each compartment by patch
no_control_plot_function(uncontrolled_epi, "ebola")
```

### Vaccination Only
We consider the case with only vaccination (no TBD).

Controls start at day `r response_time`.

```{r ebola-exp4-vacc_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_v_only, "ebola")
```

```{r ebola-exp4-vacc_only_cost, fig.show="hold", out.width = "50%"}
# plot costs
cost_plot_function(exper_v_only, "ebola")

# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_v_only, "ebola")
```

### TBD Only
We also consider the case with only TBD (no vaccination).

Controls start at day `r response_time``.

```{r ebola-exp4-TBD_only_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_u_only, "ebola")
```

```{r ebola-exp4-TBD_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")

# plot costs
cost_plot_function(exper_u_only, "ebola")
```

### Both Vaccination and TBD

Lastly, we show how results change if we optimize both controls.

Controls start at day `r response_time`.

```{r ebola-exp4-both_trajectories, fig.show="hold", out.width = "100%"}
# plot states
trajectory_plot_function(exper_both, "ebola")
```

```{r ebola-exp4-both_cost, fig.show="hold", out.width = "50%"}
# Measure the relative differences in each type of cost between unique and uniform
relcost_plot_function(exper_u_only, "ebola")
# plot costs
cost_plot_function(exper_both, "ebola")
```