---
title: "Cholera Optimal Control Results"
output: html_document
---

```{r setup, include=FALSE}
# markdown settings-------------------------------------------------------------
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# required packages-------------------------------------------------------------
library(ggplot2)
library(deSolve)
library(reshape2)
library(tidyverse)
library(cowplot)
library(pracma)
library(RColorBrewer)

# to paralellize
library(doParallel)
library(foreach)
registerDoParallel(4) # update this 4 if you want to use more cores
```

```{r oc_setup}
# load optimal control files--------------------------------------------------
source("CholeraSIRW_ODE.R")
source("Cholera_params.R")
source("Cholera_OCfunc.R")
source("Cholera_adjoints.R")
source("Cholera_analysisfunc.R")

# initial guesses - controls----------------------------------------------------
guess_v1 <- rep(0, length(times))
guess_v2 <- rep(0, length(times))
guess_u1 <- rep(0, length(times))
guess_u2 <- rep(0, length(times))

# setup baseline optimal control parameters-------------------------------------
tol = 0.01 # tolerance parameter for optimization
oc_params <- c(b1 = 1, b2 = 1, # cost of cases
               C1 = 0.125, C2 = 0.125,  # cost of vaccinations
               epsilon1  = 10000, epsilon2 = 10000, # non-linearity for vacc
               D1 = 0.125, D2 = 0.125, # cost of sanitation 
               eta1 = 100, eta2 = 100 ) # non-linearity for sanitation

# setup control_type parameters to test-----------------------------------------
# unique: optimal control can vary between patches
# uniform: optimal control must be the same in each patch
test_params <- data.frame(
  m1 = 0,
  control_type = c("unique", "uniform") # how to optimize control in each patch
)
```

```{r plot_setup}
compartment_names = c("susceptible", "infected", "recovered", "water", "vaccination", "sanitation")
names(compartment_names) = c("S", "I", "R", "W", "v", "u")
```

## Cholera Optimal Control

Here we keep the results of various numerical experiments of optimal control outcomes in the Cholera model. Specifically, we focus on the difference between the optimal strategy, when control is customized to each patch versus the "one size fits all strategy" where both patches are forced to be given the same control. 

## No control

```{r}
# read in csv with solution to ode without control------------------------------
uncontrolled_epi <- read.csv("analysis/initialOutbreak_noControl_forIC.csv")

# plot the time series in each compartment by patch ----------------------------
uncontrolled_epi %>%
  # melt from wide to long format
  melt(c("time")) %>%
  # use variable names to distinguish compartment and patch
  mutate(compartment = substr(variable, 1, 1),
         compartment = factor(compartment, levels = c("S", "I", "R", "W")),
         patch = substr(variable, 2, 2)) %>%
  # plot time series
  ggplot(aes(x = time, y = value, color = patch)) + 
  geom_line() +
  facet_wrap(vars(compartment), scales = "free") +
  theme_bw() + 
  theme(axis.title.y = element_blank(),
        legend.position = "bottom")
  
```


## Vaccination Only
Next we consider the case with only vaccination (no sanitation).

```{r vacc_only_oc}
# exclude sanitation by setting u max to 0--------------------------------------
bounds_vonly <- bounds
bounds_vonly["U1_max"] = 0
bounds_vonly["U2_max"] = 0

# calculate optimal control-----------------------------------------------------
# run OC across multiple parameters
exper_v_only <- test_mult_params(
  test_params = test_params,
  return_type = c("X", "j", "v", "u"),
  base_params = c(params, oc_params),
  guess_v1 = guess_v1, guess_v2 = guess_v2,
  guess_u1 = guess_u1, guess_u2 = guess_u2,
  IC = IC, bounds = bounds_vonly, times = times, tol = tol
)

# reformat output---------------------------------------------------------------
exper_v_only <- reformat_mult_params_output(output = exper_v_only, 
                                            test_params = test_params)
```

```{r vacc_only_plot_time_series}
# plot states 
exper_v_only$v %>% 
  left_join(exper_v_only$u) %>%
  left_join(exper_v_only$X) %>% 
  melt(c("m1", "control_type", "test_case", "time")) %>%
  mutate(patch = substr(variable, 2,2), 
         variable = substr(variable, 1,1)) %>%
  filter(control_type %in% c("uniform", "unique")) %>%
  mutate(plot_var = ifelse(control_type == "unique", paste("patch", patch), "uniform"),
         variable = factor(variable, levels = c("S", "I", "R", "W", "v", "u"))) %>%
  ggplot(aes(x = time, y = value, color = patch, linetype = control_type)) + 
  geom_line(size = 1)+
  facet_wrap(vars(variable), scales = "free", labeller = labeller(variable = compartment_names)) +
  scale_linetype_manual(values = c("dotted", "solid")) +
  theme_bw()+
  theme(legend.position = "bottom",
        panel.grid = element_blank())
```


## Sanitation Only

## Both Vaccination and Sanitation


